---
title: "Take Home Exercise 2"
author:
- name: "Yu Di"
  affiliation: "Singapore Management University"
description: |
  Remake and beautify classmates' submissions of exercise 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task Description

In this take-home exercise, you are required to:

 - select one of the Take-home Exercise 1 prepared by your classmate,
 - critic the submission in terms of clarity and aesthetics, and
remake the original design by using the data visualisation principles and best practice you had learned in Lesson.


## Load data

```{r load data, echo=TRUE}
packages = c('tidyverse', 'ggdist', 'ggridges',
             'patchwork', 'ggthemes', 'hrbrthemes',
             'ggrepel', 'ggforce')

for(p in packages) {
  if(!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
participants_data <- read_csv('data/Participants.csv')
```

I choose Huang Anni's exercise to do this exercise 2.

## Charts to Modify

The previous first charts is:

```{r}
participants_data$age_state = ifelse(
    test = participants_data$age > 30,
    yes = "Old",
    no = "Young"
)
# Make it a factor
participants_data$age_state = factor(
    participants_data$age_state,
    levels = c("Old", "Young")
)
p1 = ggplot(data=subset(participants_data,age_state=='Young'), 
       aes(x = joviality)) + 
  geom_density() +
  ggtitle("Distribution of joviality for young people")+
  theme(plot.title = element_text(size = 10))
p2 = ggplot(data=subset(participants_data,age_state=='Old'), 
       aes(x = joviality)) + 
  geom_density() +
  ggtitle("Distribution of joviality for old people")+
  theme(plot.title = element_text(size = 10))
p3 = ggplot(data=participants_data,
aes(x= joviality,
fill = age_state)) +
geom_density(alpha=0.2) +
  annotate("text", x = 0.7, y = 1.2, label = "Young people tend\n to be happier",size=3,color='#4682B4') + 
  ggtitle("Distribution of joviality in different Age Groups")+
  theme(plot.title = element_text(size = 10))
(p1 / p2) | (p3+
  scale_y_continuous(name="density", limits=c(0.0, 1.2)))
```

The left part's meaning is the same as the right one, both want to compare the joviality of different age groups. So we just simplify this graph

```{r}
ggplot(data=participants_data,
aes(x= joviality,
fill = age_state)) +
geom_density(alpha=0.2) + coord_cartesian(ylim = c(0.3, 1.2)) +
  annotate("text", x = 0.8, y = 1.15, label = "Young people are happier",size=5,color='#4682B4') +
  ggtitle("Distribution of joviality (Old: Age>30, Young: Age<30)") +
  theme_ipsum(grid = "Y", axis_title_size = 18) + 
  theme(plot.title = element_text(size = 15), axis.title.y = element_blank())
```

For the second chart

```{r}
ggplot(data=participants_data,
aes(x= joviality,
fill = haveKids)) +
geom_histogram(alpha=0.2) +
  annotate("text", x = 0.7, y = 20, label = "People with kids tend\n to be happier at the start of the study",size=3,color='red') + 
  ggtitle("Distribution of joviality with and without kids")+
  theme(plot.title = element_text(size = 10))

agg_happy <- participants_data %>% 
  select(c("householdSize","joviality")) %>%
  group_by(householdSize) %>% 
  summarise(joviality=mean(joviality))

happy_sorted <- agg_happy %>%
  arrange(desc(householdSize))
ggplot(data=happy_sorted,
aes(y = joviality,
x= householdSize)) +
geom_line(linetype = "dashed",color='black',arrow = arrow(type = "closed"))+
  geom_point(stat = "identity",
             position = "identity")+
  ggtitle("Change of average joviality when householdSize increase")+
  annotate("text", 
           x = 2.5, 
           y = 0.495, 
           label = "People tend \n to be happier\n at the start of study with\n larger householdSize",size=3,color='#4682B4') + 
  theme(plot.title = element_text(size = 12))
```

The expression is correct, but the color is not very fashion and it should not use opacity too much; The comparison of household contains only 3 groups, so it should not use line plot. Instead, barplot is a good alternative. So to modify it, we change it to: 

```{r}
ggplot(data=participants_data,
aes(x= joviality, fill = haveKids)) +
geom_histogram(binwidth = 0.05, alpha=0.5) +
  annotate("text", x = 0.6, y = 57, label = "People with kids generally tend to be happier ",size=3,color='red') + 
  ggtitle("Distribution of joviality partitioned by kids")+
  theme_ipsum(grid = "Y", axis_title_size = 18) + 
  theme(plot.title = element_text(size = 15), axis.title.y = element_blank(), legend.position = "top")

agg_happy <- participants_data %>% 
  select(c("householdSize","joviality")) %>%
  group_by(householdSize) %>% 
  summarise(joviality=mean(joviality))
agg_happy$joviality <- round(agg_happy$joviality, 3)

ggplot(data=agg_happy, aes(x=householdSize, y=joviality)) + 
  geom_bar(stat = "identity", width = 0.5, fill="steelblue") +
  coord_cartesian(ylim = c(0.45, 0.51)) + 
  ggtitle("Average Joviality for different HouseholdSize")+
  geom_text(aes(label = joviality), vjust = 1.5, colour = "white")+
  theme_ipsum(grid = "Y", axis_title_size = 18) + 
  theme(plot.title = element_text(size = 15), axis.title.y = element_blank())
```

For the third chart, 

```{r}
ggplot(participants_data, aes(x = joviality, y = educationLevel, fill = educationLevel)) +
  geom_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE,
                      quantiles = 4, 
                      quantile_lines = TRUE,
                      alpha = .2) +
  theme_ridges() + 
  scale_fill_viridis_d(name = "Quartiles")+
  ggtitle("Distribution of joviality in different Education Level")+
  theme(plot.title = element_text(size = 12))
```

the expression is correct, but the legend takes almost half of the plot, which is not very reasonable, so I just simply modify it by:

```{r}
ggplot(participants_data, aes(x = joviality, y = educationLevel, fill = educationLevel)) +
  geom_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE,
                      quantiles = 4, 
                      quantile_lines = TRUE,
                      alpha = .2) +
  theme_ridges() + 
  scale_fill_viridis_d(name = "Quartiles")+
  ggtitle("Distribution of joviality in different Education Level")+
  theme(plot.title = element_text(size = 12), legend.position = "top")
```

